Latent Predicate Networks: Concept Learning with Probabilistic Context-Sensitive Grammars

### 2014-10-21 drafts/cfp/draft.tex
Added experiments section.

@@

See Figures \ref{fig:parseexample} and \ref{fig:results}.

{\it Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec a
  diam lectus. Sed sit amet ipsum mauris. Maecenas congue ligula ac
  quam viverra nec consectetur ante hendrerit. Donec et mollis
  dolor. Praesent et diam eget libero egestas mattis sit amet vitae
  augue. Nam tincidunt congue enim, ut porta lorem lacinia
  consectetur. Donec ut libero sed arcu vehicula ultricies a non
  tortor. Lorem ipsum dolor sit amet, consectetur adipiscing
  elit. Aenean ut gravida lorem. Ut turpis felis, pulvinar a semper
  sed, adipiscing id dolor. Pellentesque auctor nisi id magna
  consequat sagittis. Curabitur dapibus enim sit amet elit pharetra
  tincidunt feugiat nisl imperdiet. Ut convallis libero in urna
  ultrices accumsan. Donec sed odio eros. Donec viverra mi quis quam
  pulvinar at malesuada arcu rhoncus. Cum sociis natoque penatibus et
  magnis dis parturient montes, nascetur ridiculus mus. In rutrum
  accumsan ultricies. Mauris vitae nisi at sem facilisis semper ac in
  est.

  Vivamus fermentum semper porta. Nunc diam velit, adipiscing ut
  tristique vitae, sagittis vel odio. Maecenas convallis ullamcorper
  ultricies. Curabitur ornare, ligula semper consectetur sagittis,
  nisi diam iaculis velit, id fringilla sem nunc vel mi. Nam dictum,
  odio nec pretium volutpat, arcu ante placerat erat, non tristique
  elit urna et turpis. Quisque mi metus, ornare sit amet fermentum et,
  tincidunt et orci. Fusce eget orci a orci congue vestibulum. Ut
  dolor diam, elementum et vestibulum eu, porttitor vel
  elit. Curabitur venenatis pulvinar tellus gravida ornare. Sed et.}

>>>>

To evaluate LPNs as a probabilistic model of concept acquisition, we
trained an LPN with $4$ latent predicates on a set of sentences
expressing successor and predessor relations in numbers between one
and ninety-nine. The training set was the collection of sentences $$X
= \{[\text{after}\, | \, \text{before}] \, \langle n \rangle \,
\text{comes} \, \langle n+1 \rangle \,|\, n \in 1,\dots,99\}.$$ The
lexicon was the set of number word corresponding to $1$ through $19$,
the decades $20, \dots, 30$ and the words ``before'' and ``after.'' It
is not difficult to manually work out an LPN that describes this
limited domain of sentences; see Figure~\ref{fig:parseexample}, for a
possible LPN derivation of an example sentence.

Although it is difficult to know how common these kinds of sentences
are in child-directed speech, words for small numbers are far more
common than words for larger ones(cite CHILDES). On the other hand,
children learning to count to large numbers rehearse the sequence. To
approximate this distribution of evidence, we drew these sample
sentences from a sum of a geometric distribution with parameter $0.5$
and a uniform distribution. These components were weighted $75\%$ and
$25\%$, respectively. We drew $2000$ examples from this distribution,
holding out the sentences in Table~\ref{tab:results} for
evaluation.

For inference, we used default $\frac{1}{D}$ pseudocounts (where $D$ is the
dimensionality of the Dirichlet distributions). We found that different random
initialization for this experiment did not lead to qualitatively different
results, though further investigation will be necessary to see how
robust our the algorithm is to local maximuma for LPNs.

We evaluated the learned model by asking for Viterbi (i.e. maximum a
posteriori) completions of the last words of each held out test
sentence. Table~\ref{tab:results} shows these completions. The grammar
correctly learns much of the structure of these sentences, including
the difference between sentences starting with ``before'' and
``after'' and the edge cases that relate decade words like ``twenty''
to non-decade words like ``twenty one.''

To inspect visually the learned grammar, we thresholded rules
according to the expected number of times they were used in parsing
the training dataset. Table~\ref{tab:grammar} shows all rules with
expected count above $1e-6$. This reduces from $2669$ to $52$ the
number of significant rules. On inspection, predicate $A_2$ forms
``before'' sentences, predicate $A_4$ forms ``after'' sentences,
predicate $A_3$ is successorship recursively defined over the decades
and ones, and predicate $A_2$ is a category for the decade words.

@@
